---
title: "XAI 3: Model-Agnostic methods"
output: html_notebook
---

## EXERCISE:

Apply PDP to the regression example of predicting bike rentals. Fit a random forest approximation for the prediction of bike rentals (**cnt**). Use the partial dependence plot to visualize the relationships the model learned. Use the slides shown in class as model.  

## QUESTION:

Analyse the influence of **days since 2011, temperature, humidity** and **wind speed** on the predicted bike counts.


```{r}
library(dplyr)
library(plotly)
library(reshape2)
library(lubridate)
library(randomForestSRC)
setwd('.')
#setwd("/Users/cmonserr/OneDrive - UPV/Trabajo_2/Asignaturas/Evaluacion de modelos/Practicas/Practica 3/Bike-Sharing-Dataset")
days <- read.csv("day.csv")
hour <- read.csv("hour.csv")

days$dteday <- as_date(days$dteday)
days_since <- select(days, workingday, holiday, temp, hum, windspeed, cnt)
days_since$days_since_2011 <- int_length(interval(ymd("2011-01-01"), days$dteday)) / (3600*24)
days_since$SUMMER <- ifelse(days$season == 3, 1, 0)
days_since$FALL <- ifelse(days$season == 4, 1, 0)
days_since$WINTER <- ifelse(days$season == 1, 1, 0)
days_since$MISTY <- ifelse(days$weathersit == 2, 1, 0)
days_since$RAIN <- ifelse(days$weathersit == 3 | days$weathersit == 4, 1, 0)
days_since$temp <- days_since$temp * 47 - 8
days_since$hum <- days_since$hum * 100
days_since$windspeed <- days_since$windspeed * 67

rf <- rfsrc(cnt~., data=days_since)

results <- select(days_since, days_since_2011, temp, hum, windspeed, cnt)
nr <- nrow(days_since)
for(c in names(results)[1:4])
{
  for(i in 1:nr){
    r <- days_since
    r[[c]] <- days_since[[c]][i]
    sal <- predict(rf, r)$predicted
    results[[c]][i] <- sum(sal) / nr
  }
}


```
Gráficos 1D:

```{r}

p1 = ggplot(days_since, aes(x=temp,y=results$temp)) + geom_line() +ylim(0,6000) + labs(x='Temperature',y='')
p2 = ggplot(days_since, aes(x=hum,y=results$hum)) +geom_line()+ylim(0,6000)+labs(x='Humidity',y='')
p3 = ggplot(days_since, aes(x=days_since_2011,y=results$days_since_2011))+geom_line()+ylim(0,6000)+ labs(x='Days since 2011',y='prediction')
p4 = ggplot(days_since, aes(x=windspeed,y=results$windspeed))+geom_line()+ylim(0,6000)+ labs(x='Wind speed',y='')
subplot(p3,p1,p2,p4,titleX = TRUE,titleY = TRUE)
```
Days_since_2011 -> A medida que pasan días desde 2011, el número de bicis que se alquilan va aumentando progresivamente. Hasta a partir del día 625 aproximadamente desde 2011 que esta tendencia cambia y empieza a disminuir el alquiler de bicis. Es decir, que a finales de 2012 el alquiler de bicis empieza a disminuir respecto a la tendencia ascendente que se observaba hasta ese momento.

Temperature -> Al igual que con la variable anterior, el alquiler de bicis aumenta conforme lo hace la temperatura, hasta que llegamos a temperaturas de alrededor de 28-30 grados donde el alquiler de bicis disminuye ligeramente. En otras palabras, cuando hace mucho frío o hace mucho calor el alquiler de bicis es menor que cuando  nos encontramos en épocas de entretiempo.

Humidity -> La humedad es una variable que no afecta al número de alquileres de bicicletas durante la mayor parte de su rango de valores. Únicamente el alquiler de bicis empieza a disminuir para valores de humedad superiores al 50%. Para valores inferiores, el número de alquileres permanece constante.

Wind Speed -> Al igual que con humidity, la velocidad del viento tiene una correlación negativa con el número de alquileres, ya que a medida que las ráfagas de viento son superiores el número de alquileres disminuye, a lo largo de todo el rango de valores de la variable 'Wind Speed'.


## EXERCISE:

Generate a 2D Partial Dependency Plot with humidity and temperature to predict the number of bikes rented depending of those parameters.

BE CAREFUL: due to the size, extract a set of random samples from the BBDD before generating the the data for the Partial Dependency Plot. 

Show the density distribution of both input features with the 2D plot as shown in the class slides. 

TIP: Use geom_tile() to generate the 2D plot. Set width and height to avoid holes. 

## QUESTION:

Interpret the results.


```{r}

sampled <- sample_n(days_since, 40)
temp <- sampled$temp
hum <- sampled$hum
th <- inner_join(data.frame(temp),data.frame(hum), by=character())
th$p <- 0

for(i in 1:nrow(th)){
  r <- days_since
  r[["temp"]] <- th[["temp"]][i]
  r[["hum"]] <- th[["hum"]][i]
  
  sal <- predict(rf, r)$predicted
  th[["p"]][i] <- sum(sal) / nr
}

```


## EXERCISE:

Apply the previous concepts to predict the **price** of a house from the database **kc_house_data.csv**. In this case, use again a random forest approximation for the prediction based on the features **bedrooms**, **bathrooms**, **sqft_living**, **sqft_lot**, **floors** and **yr_built**. 
Use the partial dependence plot to visualize the relationships the model learned.

BE CAREFUL: due to the size, extract a set of random samples from the BBDD before generating the data for the Partial Dependency Plot. 

## QUESTION:

Analyse the influence of **bedrooms, bathrooms, sqft_living** and **floors** on the predicted price.


```{r}

d <- read.csv("kc_house_data.csv")

sampled <- sample_n(d, 1000)

sampled <- select(sampled, bedrooms, bathrooms, sqft_living, sqft_lot, floors, yr_built, price)

rf <- rfsrc(price~., data=sampled)

results <- select(sampled, bedrooms, bathrooms, sqft_living, floors, price)
nr <- nrow(sampled)
for(c in names(results)[1:4])
{
  for(i in 1:nr){
    r <- sampled
    r[[c]] <- sampled[[c]][i]
    sal <- predict(rf, r)$predicted
    results[[c]][i] <- sum(sal) / nr
  }
}

```

